---
title: "Discussion_single year (ELAPSE)"
author: "Youchen"
date: "01/26/2021"
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.width = 6, fig.height = 6, warning=F)
source("scr/fun_call_lib.R")
```
```{r read data}
eu_bnd <- st_read("../expanse_shp/eu_expanse2.shp")

```



# Location
The model performance of GWR kinda vary greatly among folds.

```{r}
csv_names <- paste0('NO2_5cv_', c(2005:2012))
years <- as.list(seq(2005, 2012))
all_test <- lapply(paste0("data/workingData/", csv_names, ".csv"), read.csv)
# lapply(all_test, nrow)

show_EM <- function(all_df_i){
   gwr=all_test[[all_df_i]]$gwr
   slr=all_test[[all_df_i]]$slr
   rf=all_test[[all_df_i]]$rf
   df_all <- data.frame(slr=error_matrix(all_test[[all_df_i]]$obs, slr),
           gwr=error_matrix(all_test[[all_df_i]]$obs, gwr),
           rf=error_matrix(all_test[[all_df_i]]$obs, rf),
           yr=years[[all_df_i]])[c(1,5,7),]
   df_all$EM = row.names(df_all)
   df_all
}
em_df <- lapply(seq_along(all_test), show_EM)
em_df <- do.call(rbind, em_df)
ggplot(em_df %>% filter(EM=='rsq') %>%  gather("model", "values", -c("yr", "EM")), aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))

# ggplot(perfm)+
#       geom_boxplot(aes(x=model, y=rsq))+
#       facet_grid(df_type~., scales ='free')+
#       labs(title=years[[i]])+
#       theme(axis.title = element_text(size = 18),
#           axis.text = element_text(size = 16),
#           legend.title = element_text(size = 16),
#           legend.text = element_text(size = 16))


```


The datasets were randomly divided into training and test data, using 'type_of_st' and 'climate_zone' as stratification.


```{r fig.height=6, fig.width=10}
tmap_mode('plot')
source("scr/fun_read_data.R")
maps_l <- lapply(seq_along(years), function(yr_i){
   no2_e_09_11 <- subset_df_yrs(no2_e_all, years[[yr_i]])
# data_all <- no2_e_09_11
print(paste0("year: ", unique(no2_e_09_11$year)))
source("scr/fun_create_fold.R")
data_all1 <- create_fold(no2_e_09_11, seed)
EU_sp <- st_as_sf(data_all1, coords = c("Xcoord", "Ycoord"))
st_crs(EU_sp) <- st_crs(3035)

   map_1 <- tm_shape(EU_sp) +
      tm_dots(size = 0.05, col="nfold",   # col = "NO2",          # popup.vars for showing values
              title = paste0('region'), palette=viridis(5), style = "cat"  #,  palette=viridis(5), style=="fixed"
             )+
      tm_shape(eu_bnd)+
      tm_borders()
   map_1
})
do.call(tmap_arrange, maps_l)
```

# predictor
* Supervised Linear regression (SLR)
* Geographically weighted regression (GWR)
* Random forests (RF)

SLR and GWR:

```{r}
source("scr/fun_call_predictor.R")
pred_c_rf <- c(pred_c, "x_trun", "y_trun")
print(pred_c)
```

RF:

```{r}
print(pred_c_rf)
```



# Performance matrix
* Supervised Linear regression (SLR)
* Geographically weighted regression (GWR)
* Random forests (RF)

Training data was used for training the models using these three algorithms. Then the test data was used for evaluating model performance.

Averaged from all folds:
## Test

```{r}
perfm <- lapply(seq_along(years), function(i){
   perfm <- lapply(paste0("data/workingData/", list.files("data/workingData/", paste0("perf_m_run2_", years[[i]]))), read.csv)
   perfm <- do.call(rbind, perfm)
   perfm %>% filter(df_type=="test") %>% group_by(df_type, year, model) %>% summarise(RMSE=mean(RMSE),
                                                                                      # RRMSE=mean(RRMSE),
                                                                                      # IQR=mean(IQR),
                                                                                      # rIQR=mean(rIQR),
                                                                                      # MAE=mean(MAE),
                                                                                      # rMAE=mean(rMAE),
                                                                                      # rsq_sd=sd(rsq),
                                                                                      rsq=mean(rsq)) #,explained_var=mean(explained_var)
})
do.call(rbind, perfm)
```

## Train

```{r}
perfm <- lapply(seq_along(years), function(i){
   perfm <- lapply(paste0("data/workingData/", list.files("data/workingData/", paste0("perf_m_run2_", years[[i]]))), read.csv)
   perfm <- do.call(rbind, perfm)
   perfm %>% filter(df_type=="train") %>% group_by(df_type, year, model) %>% summarise(RMSE=mean(RMSE),
                                                                                      # RRMSE=mean(RRMSE),
                                                                                      # IQR=mean(IQR),
                                                                                      # rIQR=mean(rIQR),
                                                                                      # MAE=mean(MAE),
                                                                                      # rMAE=mean(rMAE),
                                                                                      # rsq_sd=sd(rsq),
                                                                                      rsq=mean(rsq)) #,explained_var=mean(explained_var)
})
do.call(rbind, perfm)
```



