---
title: "Discussion_single year (ELAPSE)"
author: "Youchen"
date: "01/26/2021"
output: html_document
  # slidy_presentation: default
  # ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.width = 6, fig.height = 6, warning=F)
source("scr/fun_call_lib.R")
library(kableExtra)
```
```{r read data}
eu_bnd <- st_read("../expanse_shp/eu_expanse2.shp")

```

```{r read prediction results (5-fold test data)}
# Here we read the predictions for test data from all five-folds
csv_names <- paste0('NO2_5cv_', c(2005:2012))
years <- as.list(seq(2005, 2012))
all_test <- lapply(paste0("data/workingData/", csv_names, ".csv"), read.csv)
# lapply(all_test, nrow)
```

The predictors used in SLR, GWR, RF were the same.

```{r predictors in LUR}
source("scr/fun_call_predictor.R")
source("scr/fun_read_data.R")
pred_c_rf <- c(pred_c) #"x_trun", "y_trun"
x_varname = names(subset_df_yrs(no2_e_all, years[[1]]) %>% dplyr::select(pred_c_rf))
print("Predictors used:")
print(x_varname)


```



# Location

The datasets were randomly divided into training and test data, using 'type_of_st' and 'climate_zone' as stratification.


```{r fig.height=6, fig.width=10}
tmap_mode('plot')
source("scr/fun_read_data.R")
maps_l <- lapply(seq_along(years), function(yr_i){
   no2_e_09_11 <- subset_df_yrs(no2_e_all, years[[yr_i]])
   # data_all <- no2_e_09_11
   print(paste0("year: ", unique(no2_e_09_11$year)))
   source("scr/fun_create_fold.R")
   data_all1 <- create_fold(no2_e_09_11, seed)
   EU_sp <- st_as_sf(data_all1, coords = c("Xcoord", "Ycoord"))
   st_crs(EU_sp) <- st_crs(3035)
   
   map_1 <- tm_shape(EU_sp) +
      tm_dots(size = 0.05, col="nfold",   # col = "NO2",          # popup.vars for showing values
              title = paste0('nfold in year ', unique(no2_e_09_11$year)), palette=viridis(5), style = "cat"  #,  palette=viridis(5), style=="fixed"
      )+
      tm_shape(eu_bnd)+
      tm_borders()
   map_1
})
do.call(tmap_arrange, maps_l)
```

# Five-fold cross-validation


```{r}


show_EM <- function(all_df_i){
   gwr=all_test[[all_df_i]]$gwr
   gwr_rf=all_test[[all_df_i]]$gwr_rf
   slr=all_test[[all_df_i]]$slr
   rf=all_test[[all_df_i]]$rf
   df_all <- data.frame(slr=error_matrix(all_test[[all_df_i]]$obs, slr),
           gwr=error_matrix(all_test[[all_df_i]]$obs, gwr),
           gwr_rf=error_matrix(all_test[[all_df_i]]$obs, gwr_rf),
           rf=error_matrix(all_test[[all_df_i]]$obs, rf),
           yr=years[[all_df_i]])[c(1,5,7),]
   df_all$EM = row.names(df_all)
   df_all
}
em_df <- lapply(seq_along(all_test), show_EM)
em_df <- do.call(rbind, em_df)
ggplot(em_df %>% filter(EM=='rsq') %>%  gather("model", "values", -c("yr", "EM")), aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))

# ggplot(perfm)+
#       geom_boxplot(aes(x=model, y=rsq))+
#       facet_grid(df_type~., scales ='free')+
#       labs(title=years[[i]])+
#       theme(axis.title = element_text(size = 18),
#           axis.text = element_text(size = 16),
#           legend.title = element_text(size = 16),
#           legend.text = element_text(size = 16))


```


```{r selected predictors in slr, fig.width = 10, fig.height = 12}
files <- list.files("data/workingData/", "SLR_summary_model_run2_2")[grepl("fold", list.files("data/workingData/", "SLR_summary_model_run2_2"))]
comp_rf_slr <- function(csv_i){
   slr_name <- files[csv_i]
   yr <- strsplit(slr_name, "_")[[1]][5] %>% as.numeric()
   nfold <- (strsplit(slr_name, "_")[[1]][7] %>% strsplit(., ".csv"))[[1]] %>% as.numeric()
   # rf_name <- paste0("RF_vi_o_", yr, "_fold_", nfold, ".csv")
   slr_result <- read.csv(paste0("data/workingData/", slr_name), header=T)
   slr_result$yr <- yr
   slr_result$nfold <- nfold
   slr_result$increR2[1]=0
   slr_result$incredR2 <- c(0, diff(slr_result$increR2))
   p1 <- ggplot(slr_result)+
      geom_col(aes(x=reorder(variables, incredR2), y=incredR2), 
               position = "dodge2", fill='khaki')+
      coord_flip() +
      theme_light()+
      labs(x = 'variable', y = 'increased R2',
           title = paste0(yr, "_fold_", nfold))+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
      theme(axis.title = element_text(size = 18),
            axis.text = element_text(size = 16),
            legend.title = element_text(size = 16),
            legend.text = element_text(size = 16),
            strip.text.y = element_text(size = 15))
   # rf_vi <- read.csv(paste0("data/workingData/", rf_name))
   source("scr/fun_plot_rf_vi.R")
   p2 <- plot_rf_vi(paste0("run2_", yr, "_fold_", nfold), var_no = 15)
   # grid.arrange(p1, p2, nrow=1)
   
   var_imp <- read.csv(paste0("data/workingData/GWR_rf_vi_run2_", yr, "_fold_", nfold, ".csv"))
   p3 <- ggplot(var_imp %>% top_n(20, vi))+
         geom_col(aes(reorder(var_name, vi), vi),
                  position = 'dodge', fill='khaki')+
         coord_flip() +
         theme_light()+
         labs(x = 'variable', y = 'importance value (impurity)',
              title = paste0(yr, "_fold_", nfold, "_gwr_rf"))+
         theme(axis.title = element_text(size = 13),
               axis.text = element_text(size = 13),
               legend.title = element_text(size = 13),
               legend.text = element_text(size = 13),
               strip.text.y = element_text(size = 12))
   grid.arrange(p1, p2, p3, nrow=2, ncol=2)
}

# comp_rf_slr(4)
pp <- lapply(seq_along(files), comp_rf_slr)
```


# Performance matrix

* Supervised Linear regression (SLR)
* Geographically weighted regression (GWR)
* Random forests (RF)

Training data was used for training the models using these three algorithms. Then the test data was used for evaluating model performance.

Averaged error matrix from all folds:

## Test
```{r}
perfm <- lapply(paste0("data/workingData/", list.files("data/workingData/", paste0("perf_m_run2_"))), read.csv)
perfm <- do.call(rbind, perfm)
perfmm <- perfm %>% filter(df_type=="test") %>% mutate(run=substr(csv_name, 1, 10), 
                            nfold=substr(csv_name, 12, 17))
ggplot(perfmm, aes(x=run, y=rsq, fill=nfold, group=nfold))+
    geom_bar(stat="identity", position = "dodge2")+
      facet_grid(model~., scales ='free')+
      # labs(title=years[[i]])+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))
```

```{r, echo=FALSE}
perfm <- lapply(seq_along(years), function(i){
   perfm <- lapply(paste0("data/workingData/", list.files("data/workingData/", paste0("perf_m_run2_", years[[i]]))), read.csv)
   perfm <- do.call(rbind, perfm)
   perfm %>% filter(df_type=="test") %>% group_by(df_type, year, model) %>% summarise(RMSE=mean(RMSE),
                                                                                      # RRMSE=mean(RRMSE),
                                                                                      # IQR=mean(IQR),
                                                                                      # rIQR=mean(rIQR),
                                                                                      # MAE=mean(MAE),
                                                                                      # rMAE=mean(rMAE),
                                                                                      # rsq_sd=sd(rsq),
                                                                                      rsq=mean(rsq))#,explained_var=mean(explained_var)
})
do.call(rbind, perfm) %>% kable(full_width=F)

```




## Train

```{r, echo=F}
perfm <- lapply(seq_along(years), function(i){
   perfm <- lapply(paste0("data/workingData/", list.files("data/workingData/", paste0("perf_m_run2_", years[[i]]))), read.csv)
   perfm <- do.call(rbind, perfm)
   perfm %>% filter(df_type=="train") %>% group_by(df_type, year, model) %>% summarise(RMSE=mean(RMSE),
                                                                                      # RRMSE=mean(RRMSE),
                                                                                      # IQR=mean(IQR),
                                                                                      # rIQR=mean(rIQR),
                                                                                      # MAE=mean(MAE),
                                                                                      # rMAE=mean(rMAE),
                                                                                      # rsq_sd=sd(rsq),
                                                                                      rsq=mean(rsq))
})
do.call(rbind, perfm) %>% kable(full_width=F)
```



