---
title: "Discussion_single year (ELAPSE)"
author: "Youchen"
date: "01/26/2021"
output: html_document
  # slidy_presentation: default
  # ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.width = 8, fig.height = 5, warning=F)
source("scr/fun_call_lib.R")
library(kableExtra)
library(gtools)
```
```{r read data}
eu_bnd <- st_read("../expanse_shp/eu_expanse2.shp")
met_str <- 'precip|temp|wind|pressure'
years <- 2000:2019
nfold=5
```

```{r useful functions}
read_perfm <- function(poll){
   csv_names <- gsub('SLR_result_all_', '', 
                     list.files('data/workingData/', 
                                paste0('SLR_result_all_o3_', poll))) %>% 
      strsplit(., '_fold_') %>% lapply(., `[[`, 1) %>% unlist() %>% unique
   years <- csv_names %>% substr(., nchar(csv_names)-3, nchar(csv_names)) %>% 
      as.numeric() %>% as.list()
   
   all_test <- lapply(paste0("data/workingData/5cv_", csv_names, ".csv"), read.csv)
   all_test
}
show_EM <- function(all_df_i){
   gwr=all_test[[all_df_i]]$gwr
   # gwr_rf=all_test[[all_df_i]]$gwr_rf
   slr=all_test[[all_df_i]]$slr
   rf=all_test[[all_df_i]]$rf
   df_all <- data.frame(slr=error_matrix(all_test[[all_df_i]]$obs, slr),
                        gwr=error_matrix(all_test[[all_df_i]]$obs, gwr),
                        # gwr_rf=error_matrix(all_test[[all_df_i]]$obs, gwr_rf),
                        rf=error_matrix(all_test[[all_df_i]]$obs, rf),
                        yr=all_test[[all_df_i]]$year)[c(1,5,7),]
   df_all$EM = row.names(df_all)
   df_all
}
comp_rf_slr <- function(csv_i, poll){
   files <- list.files("data/workingData/", paste0("SLR_summary_model_o3_", poll))[grepl("fold", list.files("data/workingData/", paste0("SLR_summary_model_o3_", poll)))]
   slr_name <- files[csv_i]
   csv_name <- gsub('.csv', '', substr(slr_name, 19, nchar(slr_name)))
   yr <- strsplit(slr_name, "_")[[1]][6]
   nfold <- (strsplit(slr_name, "_")[[1]][8] %>% strsplit(., ".csv"))[[1]] %>% as.numeric()
   # rf_name <- paste0("RF_vi_o_", yr, "_fold_", nfold, ".csv")
   slr_result <- read.csv(paste0("data/workingData/", slr_name), header=T)
   slr_result$yr <- yr
   slr_result$nfold <- nfold
   slr_result$increR2[1]=0
   slr_result$incredR2 <- c(0, diff(slr_result$increR2))
   p1 <- ggplot(slr_result)+
      geom_col(aes(x=reorder(variables, incredR2), y=incredR2), 
               position = "dodge2", fill='khaki')+
      coord_flip() +
      theme_light()+
      labs(x = 'variable', y = 'increased R2',
           title = paste0(yr, "_fold_", nfold))+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
      theme(axis.title = element_text(size = 11),
            axis.text = element_text(size = 11),
            legend.title = element_text(size = 11),
            legend.text = element_text(size = 11),
            strip.text.y = element_text(size = 11))
   # rf_vi <- read.csv(paste0("data/workingData/", rf_name))
   source("../EXPANSE_algorithm/scr/fun_plot_rf_vi.R")
   p2 <- plot_rf_vi(csv_name, var_no = 20)
   grid.arrange(p1, p2, nrow=1, ncol=2)
}

extract_slr_var <- function(csv_i, poll){
   files <- list.files("data/workingData/", paste0("SLR_summary_model_o3_", poll))[grepl("fold", list.files("data/workingData/", paste0("SLR_summary_model_o3_", poll)))]
   slr_name <- files[csv_i]
   
   yr <- strsplit(slr_name, "_")[[1]][6]
   nfold <- (strsplit(slr_name, "_")[[1]][8] %>% strsplit(., ".csv"))[[1]] %>% as.numeric()
   # rf_name <- paste0("RF_vi_o_", yr, "_fold_", nfold, ".csv")
   slr_result <- read.csv(paste0("data/workingData/", slr_name), header=T)
   slr_result$yr <- yr
   slr_result$nfold <- nfold
   slr_result$increR2[1]=0
   slr_result$incredR2 <- c(0, diff(slr_result$increR2))
   slr_var <- slr_result$variables[-1]
   if(any(grepl(met_str, slr_var))){
      slr_var[grepl(met_str, slr_var)] <- gsub('(\\_\\d+).*', '', slr_var[grepl(met_str, slr_var)])  
   }
   data.frame(var_name=slr_var, yr=yr, nfold=nfold)
}
extract_rf_var <- function(csv_i, poll){
   files <- list.files("data/workingData/", paste0("SLR_summary_model_o3_", poll))[grepl("fold", list.files("data/workingData/", paste0("SLR_summary_model_o3_", poll)))]
   slr_name <- files[csv_i]
   csv_name <- gsub('.csv', '', substr(slr_name, 19, nchar(slr_name)))
   yr <- strsplit(slr_name, "_")[[1]][6]
   nfold <- (strsplit(slr_name, "_")[[1]][8] %>% strsplit(., ".csv"))[[1]] %>% as.numeric()
   
   var_importance <- read.csv(paste0('data/workingData/RF_vi_', csv_name, '.csv'), 
                              header = T)
   rf_var10 <- (var_importance %>% top_n(10, vi))$var_name
   
   if(any(grepl(met_str, rf_var10))){
      rf_var10[grepl(met_str, rf_var10)] <- gsub('(\\_\\d+).*', '', rf_var10[grepl(met_str, rf_var10)])  
   }
   data.frame(var_name=rf_var10, yr=yr, nfold=nfold)
}
create_heatmap <- function(poll){
   files <- list.files("data/workingData/", paste0("SLR_summary_model_o3_", poll))[grepl("fold", list.files("data/workingData/", paste0("SLR_summary_model_o3_", poll)))]
   slr_vars <- lapply(seq_along(files), extract_slr_var, poll=poll) %>% do.call(rbind, .)
   rf_vars <- lapply(seq_along(files), extract_rf_var, poll=poll) %>% do.call(rbind, .)
   slr_tbl <- with(slr_vars, table(var_name, yr))
   rf_tbl <- with(rf_vars, table(var_name, yr))
   slr_tbl <- slr_tbl[mixedorder(row.names(slr_tbl), decreasing=T),]
   rf_tbl <- rf_tbl[mixedorder(row.names(rf_tbl), decreasing=T),]
   print(heatmap(rf_tbl, Colv = NA, Rowv = NA, scale="column", main=paste0('RF: ', poll)))
   print(heatmap(slr_tbl, Colv = NA, Rowv = NA, scale="column", main=paste0('SLR: ', poll)))
}
```

# PM2.5

```{r read prediction results (5-fold test data)}
# Here we read the predictions for test data from all five-folds
target_poll = 'PM2.5'
all_test <- read_perfm(target_poll)

```


The predictors used in SLR, GWR, RF were the same.

```{r predictors in LUR}
source("../expanse_multiyear/src/00_fun_read_data_gee.R")
no2_e_09_11 <- read_data(target_poll, 2010)
source("../expanse_multiyear/src/00_fun_create_fold.R")
data_all1 <- create_fold(no2_e_09_11, seed, strt_group=c("sta_type", "zoneID"), 
                               nfold = nfold)
#f# SLR: select predictors
# source("scr/o_00_01_call_predictor.R")
exc_names <- c("sta_code", "component_code", "component_caption", "obs",
               'system.index', '.geo', 
               "year", "id", "country_name", "sta_type", "area_type", "areaid", 
               "index", "nfold", "xcoord", "ycoord", 'cntr_code')  #zoneID
pred_c <- names(data_all1)[!names(data_all1)%in%exc_names]
# pred_c
```



# Five-fold cross-validation

The model performance of GWR kinda vary greatly among folds.

```{r}

em_df <- lapply(seq_along(all_test), show_EM)
em_df <- do.call(rbind, em_df)
ggplot(em_df %>% filter(EM=='rsq') %>%  gather("model", "values", -c("yr", "EM")), aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
   labs(y="R squared")+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))+
   geom_hline(aes(yintercept=0.587))+
   geom_hline(aes(yintercept=0.664))

ggplot(em_df %>% filter(EM=='RMSE') %>%  gather("model", "values", -c("yr", "EM")), 
       aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
   labs(y="RMSE")+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))+
   geom_hline(aes(yintercept=3.3))+
   geom_hline(aes(yintercept=2.97))
```


```{r selected predictors in slr, fig.width = 10, fig.height = 3.6}

# comp_rf_slr(4)
# lapply(seq_along(files), comp_rf_slr)


```

```{r, fig.width = 8, fig.height = 6}
create_heatmap(target_poll)
```


# PM10


```{r read1 prediction results (5-fold test data) pm10}
# Here we read the predictions for test data from all five-folds
target_poll = 'PM10'
all_test <- read_perfm(target_poll)
```


The predictors used in SLR, GWR, RF were the same.

```{r  }
source("../expanse_multiyear/src/00_fun_read_data_gee.R")
no2_e_09_11 <- read_data(target_poll, 2010)
source("../expanse_multiyear/src/00_fun_create_fold.R")
data_all1 <- create_fold(no2_e_09_11, seed, strt_group=c("sta_type", "zoneID"), 
                               nfold = nfold)
#f# SLR: select predictors
# source("scr/o_00_01_call_predictor.R")
exc_names <- c("sta_code", "component_code", "component_caption", "obs", 
               'system.index', '.geo',
               "year", "id", "country_name", "sta_type", "area_type", "areaid", 
               "index", "nfold", "xcoord", "ycoord", 'cntr_code')  #zoneID
pred_c <- names(data_all1)[!names(data_all1)%in%exc_names]
# pred_c

```




### Five-fold cross-validation

The model performance of GWR kinda vary greatly among folds.

```{r}
em_df <- lapply(seq_along(all_test), show_EM)
em_df <- do.call(rbind, em_df)
ggplot(em_df %>% filter(EM=='rsq') %>%  gather("model", "values", -c("yr", "EM")), aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
   labs(y="R squared")+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))

ggplot(em_df %>% filter(EM=='RMSE') %>%  gather("model", "values", -c("yr", "EM")), 
       aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
   labs(y="RMSE")+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))


```


```{r  , fig.width = 10, fig.height = 3.6}
# files <- list.files("data/workingData/", paste0("SLR_summary_model_o3_", target_poll))[grepl("fold", list.files("data/workingData/", paste0("SLR_summary_model_o3_", target_poll)))]

# # comp_rf_slr(4)
# lapply(seq_along(files), comp_rf_slr)

```

### Heatmap of top-10 important variables in RF

```{r, fig.width = 8, fig.height = 6}
create_heatmap(target_poll)
```

# NO2


```{r read1 prediction results (5-fold test data) no2}
# Here we read the predictions for test data from all five-folds
target_poll = 'NO2'
all_test <- read_perfm(target_poll)
```


The predictors used in SLR, GWR, RF were the same.

```{r  }
source("../expanse_multiyear/src/00_fun_read_data_gee.R")
no2_e_09_11 <- read_data(target_poll, 2010)
source("../expanse_multiyear/src/00_fun_create_fold.R")
data_all1 <- create_fold(no2_e_09_11, seed, strt_group=c("sta_type", "zoneID"), 
                               nfold = nfold)
#f# SLR: select predictors
# source("scr/o_00_01_call_predictor.R")
exc_names <- c("sta_code", "component_code", "component_caption", "obs", 
               'system.index', '.geo',
               "year", "id", "country_name", "sta_type", "area_type", "areaid", 
               "index", "nfold", "xcoord", "ycoord", 'cntr_code')  #zoneID
pred_c <- names(data_all1)[!names(data_all1)%in%exc_names]
# pred_c

```




# Five-fold cross-validation

The model performance of GWR kinda vary greatly among folds.

```{r}
em_df <- lapply(seq_along(all_test), show_EM)
em_df <- do.call(rbind, em_df)
ggplot(em_df %>% filter(EM=='rsq') %>%  gather("model", "values", -c("yr", "EM")), aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
   labs(y="R squared")+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))+
   geom_hline(aes(yintercept=0.575))

ggplot(em_df %>% filter(EM=='RMSE') %>%  gather("model", "values", -c("yr", "EM")), 
       aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
   labs(y="RMSE")+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))+
   geom_hline(aes(yintercept=09.51))


```


```{r  , fig.width = 10, fig.height = 3.6}
# files <- list.files("data/workingData/", paste0("SLR_summary_model_o3_", target_poll))[grepl("fold", list.files("data/workingData/", paste0("SLR_summary_model_o3_", target_poll)))]

# # comp_rf_slr(4)
# lapply(seq_along(files), comp_rf_slr)

```

### Heatmap of top-10 important variables in RF

```{r, fig.width = 8, fig.height = 6}
create_heatmap(target_poll)
```




```{r, echo=F}
# ## Performance matrix
# 
# * Supervised Linear regression (SLR)
# * Geographically weighted regression (GWR)
# * Random forests (RF)
# 
# Training data was used for training the models using these three algorithms. Then the test data was used for evaluating model performance.
# 
# Averaged error matrix from all folds:
# 
# ### Test
# perfm <- lapply(paste0("data/workingData/", list.files("data/workingData/", paste0("perf_m_o3_", target_poll, "_"))), read.csv)
# perfm <- do.call(rbind, perfm)
# perfmm <- perfm %>% filter(df_type=="test") %>% mutate(n_fold=as.factor(n_fold))
# ggplot(perfmm, aes(x=year, y=rsq, fill=n_fold, group=n_fold))+
#     geom_bar(stat="identity", position = "dodge2")+
#       facet_grid(model~., scales ='free')+
#       # labs(title=years[[i]])+
#       theme(axis.title = element_text(size = 12),
#           axis.text = element_text(size = 11),
#           legend.title = element_text(size = 11),
#           legend.text = element_text(size = 11),
#           strip.text.y = element_text(size = 11))
```




```{r, echo=F}

# ## Train
# perfm <- lapply(paste0("data/workingData/", list.files("data/workingData/", paste0("perf_m_o3_", target_poll, "_"))), read.csv)
# perfm <- do.call(rbind, perfm)
# perfmm <- perfm %>% filter(df_type=="train") %>% mutate(n_fold=as.factor(n_fold))
# ggplot(perfmm, aes(x=year, y=rsq, fill=n_fold, group=n_fold))+
#     geom_bar(stat="identity", position = "dodge2")+
#       facet_grid(model~., scales ='free')+
#       # labs(title=years[[i]])+
#       theme(axis.title = element_text(size = 12),
#           axis.text = element_text(size = 11),
#           legend.title = element_text(size = 11),
#           legend.text = element_text(size = 11),
#           strip.text.y = element_text(size = 11))
```


```{r fig.height=6, fig.width=10}

# # Location
# 
# The datasets were randomly divided into training and test data, using 'type_of_st' and 'climate_zone' as stratification.

# tmap_mode('plot')
# # tmap_mode('view')
# maps_l <- lapply(seq_along(years), function(yr_i){
#    no2_e_09_11 <- read_data(target_poll, years[[yr_i]])
#    # data_all <- no2_e_09_11
#    print(paste0("year: ", unique(no2_e_09_11$year)))
#    source("../expanse_multiyear/src/00_fun_create_fold.R")
# data_all1 <- create_fold(no2_e_09_11, seed, strt_group=c("sta_type", "zoneID"), 
#                                nfold = nfold)
#    EU_sp <- st_as_sf(data_all1, coords = c("xcoord", "ycoord"))
#    st_crs(EU_sp) <- st_crs(3035)
#    
#    map_1 <- tm_shape(EU_sp) +
#       tm_dots(size = 0.05, col="nfold",   # col = "NO2",          # popup.vars for showing values
#               title = paste0('nfold in year ', unique(no2_e_09_11$year)), palette=viridis(5), style = "cat"  #,  palette=viridis(5), style=="fixed"
#       )+
#       tm_shape(eu_bnd)+
#       tm_borders()
#    map_1
# })
# # do.call(tmap_arrange, maps_l)
# maps_l
```

# O3


```{r read1 prediction results (5-fold test data) o3}
# Here we read the predictions for test data from all five-folds
target_poll = 'O3'
all_test <- read_perfm(target_poll)
```


The predictors used in SLR, GWR, RF were the same.

```{r  }
source("../expanse_multiyear/src/00_fun_read_data_gee.R")
no2_e_09_11 <- read_data(target_poll, 2010)
source("../expanse_multiyear/src/00_fun_create_fold.R")
data_all1 <- create_fold(no2_e_09_11, seed, strt_group=c("sta_type", "zoneID"), 
                               nfold = nfold)
#f# SLR: select predictors
# source("scr/o_00_01_call_predictor.R")
exc_names <- c("sta_code", "component_code", "component_caption", "obs", 
               'system.index', '.geo',
               "year", "id", "country_name", "sta_type", "area_type", "areaid", 
               "index", "nfold", "xcoord", "ycoord", 'cntr_code')  #zoneID
pred_c <- names(data_all1)[!names(data_all1)%in%exc_names]
# pred_c

```




### Five-fold cross-validation

The model performance of GWR kinda vary greatly among folds.

```{r}
em_df <- lapply(seq_along(all_test), show_EM)
em_df <- do.call(rbind, em_df)
ggplot(em_df %>% filter(EM=='rsq') %>%  gather("model", "values", -c("yr", "EM")), aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
   labs(y="R squared")+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))

ggplot(em_df %>% filter(EM=='RMSE') %>%  gather("model", "values", -c("yr", "EM")), 
       aes(x=yr, y=values, fill=model))+
      geom_bar(stat="identity", position = "dodge2")+
      # facet_grid(EM~., scales ='free')+
      # labs(title=years[[i]])+
   labs(y="RMSE")+
      theme(axis.title = element_text(size = 18),
          axis.text = element_text(size = 16),
          legend.title = element_text(size = 16),
          legend.text = element_text(size = 16),
          strip.text.y = element_text(size = 15))


```


```{r  , fig.width = 10, fig.height = 3.6}
# files <- list.files("data/workingData/", paste0("SLR_summary_model_o3_", target_poll))[grepl("fold", list.files("data/workingData/", paste0("SLR_summary_model_o3_", target_poll)))]

# # comp_rf_slr(4)
# lapply(seq_along(files), comp_rf_slr)

```

### Heatmap of top-10 important variables in RF

```{r, fig.width = 8, fig.height = 6}
create_heatmap(target_poll)
```
