---
title: "Discussion_multiple-year modelling (ELAPSE) LME test"
author: "Youchen"
date: "02/11/2021"
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.width = 6, fig.height = 6, warning=F)
library(dplyr)
library(APMtools)
library(lme4) # linear mixed effect models
library(CAST) # For dividing training and test data (CreateSpacetimeFolds)
library(kableExtra)
```

1. Years and station IDs were included as random effects for the intercept for each step of selecting a candidate predictor.
*lmer(POLL~pred[,i] + (1|stations) + (1|years))* 

2. only look at the fixed effect for R2, p-value and VIF for the exclusion step

Different periods of years were tried: 

* 2009-2011 
* 2008-2012
* 2006-2010
* 2006-2012

(Predictor values remain the same for different years)

```{r}
library(dplyr)
library(APMtools)
library(lme4) # linear mixed effect models
library(CAST) # For dividing training and test data (CreateSpacetimeFolds)
seed <- 123

## Read in data (elapse NO2 2010 with climate zones included)
elapse_no2 <- read.csv("../EXPANSE_predictor/data/processed/no2_2010_elapse_climate.csv",
                       encoding = "utf-8")
## Read in data (airbase observations 1990s-2012)
no2 <- read.csv("../airbase/EXPANSE_APM/data/processed/ab_v8_yr_no2.csv")
# rename data
elapse_no2 <- dplyr::rename(elapse_no2, station_european_code=ï..Station)
# reduce airbase data
no2 <- no2 %>% dplyr::rename(year=statistics_year, obs=statistic_value)
## subset stations that are included in the elapse (cause at this stage, we don't have the predictor maps...)
no2_e <- no2 %>% filter(no2$station_european_code%in%unique(elapse_no2$station_european_code))
no2_e_all <- left_join(no2_e, elapse_no2, by="station_european_code")

## subset samples (for multiple years or each year)
subset_df_yrs <- function(obs_df, yr_target){
   no2_e_sub <- obs_df %>% filter(year%in%yr_target)
   no2_e_sub
}
csv_names <- paste0('run2_', c('09-11', '08-12', '06-10', '06-12'))
years <- list(2009:2011, 2008:2012, 2006:2010, 2006:2012)
lme_result <- lapply(paste0("data/workingData/LME_summary_model_", csv_names, ".csv") , read.csv)
slr_result <- lapply(paste0("data/workingData/SLR_summary_model_", csv_names, ".csv") , read.csv)

out_pred <- function(i, model_result, modelname){
   data.frame(pred=model_result[[i]]$variables[-1], 
              period=paste0(substr(csv_names[[i]], 12, nchar(csv_names[[i]])),
                            "_", modelname))
}
slr_pred <- lapply(seq_along(years), out_pred, model_result=slr_result, "slr")
lme_pred <- lapply(seq_along(years), out_pred, model_result=lme_result, "lme")
slr_pred <- do.call(rbind, slr_pred)
lme_pred <- do.call(rbind, lme_pred)
# table(rbind(slr_pred, lme_pred))
tbl_comp <- table(rbind(slr_pred, lme_pred))
tbl_comp <- tbl_comp %>% apply(., 1, sum) %>% cbind(tbl_comp,total=.)
tbl_comp <- tbl_comp[order(tbl_comp[,"total"], row.names(tbl_comp), decreasing = T), ]

tbl1 <- tbl_comp %>% kbl(full_width = F)

for (i in seq_along(unique(tbl_comp[,"total"]))) {
  tbl1 <- row_spec(tbl1, which(tbl_comp[,"total"]==unique(tbl_comp[,"total"])[i]) %>% as.numeric(), background = spec_color(unique(tbl_comp[,"total"]) %>% as.numeric())[i])
}
tbl1
   # row_spec(row= 1:nrow(tbl_comp), col=spec_color(tbl_comp[,"total"] %>% as.numeric()))


# tbl_comp_a <- as.data.frame(tbl_comp) %>% mutate(model=substr(period, 7, 10)) %>% 
#    group_by(pred, model) %>% 
#    summarise(Freq=sum(Freq)) %>% 
#    arrange(desc(Freq), pred) #%>%
#    # mutate(tmp=which(duplicated(pred))) %>% 
#    # arrange(desc(Freq), pred, tmp)
# tbl_comp_a
# tbl_comp_a[c(which(duplicated(tbl_comp_a$pred)&(Freq=4)), which(duplicated(tbl_comp_a$pred)&(Freq=4))-1),] %>% 
#    mutate(tmp=T) %>% 
#    rbind(mutate(tbl_comp_a[-c(which(duplicated(tbl_comp_a$pred)), which(duplicated(tbl_comp_a$pred))-1),],
#                 tmp=F)) %>% 
#    arrange(desc(Freq), desc(tmp), pred) %>% View
```

# performance matrix



```{r}
tmp <- matrix(nrow=length(csv_names)*2, ncol=3)
tmp_rowname <- vector("character", length=length(csv_names)*2)
for(i in seq_along(years)){
   
   csv_name <- csv_names[i]
   csv_name
   no2_e_09_11 <- subset_df_yrs(no2_e_all, years[[i]])
   data_all <- no2_e_09_11
   #f# subset cross-validation data (5-fold cross-validation)
   #f# stratified by station types, climate zones and/or years
   set.seed(seed)
   data_all$index <- 1:nrow(data_all)
   # Test only leave location out first 
   # Method 1: (easier to use)
   folds=CreateSpacetimeFolds(
      data_all,
      spacevar = "station_european_code",     # leave location out
      timevar = NA,
      k = 5,
      class = NA,
      seed = seed
   )
   train_sub <- data_all[folds$index[[1]], ]
   test_sub <- data_all[folds$indexOut[[1]], ]
   source("scr/fun_call_predictor.R")
   #f# SLR: define/preprocess predictors (direction of effect)
   source("scr/fun_slr_proc_in_data.R")
   train_sub <- proc_in_data(train_sub, neg_pred)
   test_sub <- proc_in_data(test_sub, neg_pred)
   
   eq_lme <- as.formula(paste0('obs~',
                               paste(lme_result[[i]]$variables[-1], collapse = "+"),
                               "+ (1|station_european_code) + (1|year)"))
   lme_model <- lmer(eq_lme, train_sub)
   
   eq_slr <- as.formula(paste0('obs~',
                               paste(slr_result[[i]]$variables[-1], collapse = "+")))
   slr_model <- lm(eq_slr, train_sub)
   
   data_all2 <- rbind(train_sub %>% mutate(df_type="train"),
                      test_sub %>% mutate(df_type='test'))
   
   
   # the prediction will use the unconditional (population-level) values for data with previously unobserved levels (or NAs).
   source("scr/fun_gen_pred_lme.R")
   lme_prediction <- gen_pred_lme(lme_model, data_all2, "obs")
   source("scr/fun_gen_pred_df.R")
   slr_prediction <- gen_pred_df(slr_model, data_all2, "obs")
   
   tb1 <- error_matrix(with(lme_prediction, obs[df_type=='test']), with(lme_prediction, lme[df_type=='test']))[c(1,5,7)]
   # print(paste0("slr: test data ", substr(csv_names[[i]], 12, nchar(csv_names[[i]]))))
   tb2 <- error_matrix(with(slr_prediction, obs[df_type=='test']), with(slr_prediction, slr[df_type=='test']))[c(1,5,7)] 
   tmp[seq(1,by=2, length.out = length(csv_names))[i],] <- tb1
   tmp_rowname[seq(1,by=2, length.out = length(csv_names))[i]] <- paste0(substr(csv_names[i], 12, nchar(csv_names[i])), "_lme")
   tmp[seq(2,by=2, length.out = length(csv_names))[i],] <- tb2
   tmp_rowname[seq(2,by=2, length.out = length(csv_names))[i]] <- paste0(substr(csv_names[i], 12, nchar(csv_names[i])), "_slr")
   colnames(tmp) <- names(tb2)
}
rownames(tmp) <- tmp_rowname
tbl1 <- tmp %>% kbl

for (i in seq_along(1:(nrow(tmp)/2))) {
  tbl1 <- row_spec(tbl1, ((i-1)*2+1):(i*2), background = spec_color((nrow(tmp)/2):1)[i])
}
tbl1
```
# Compare predictions from different models
* The year 2009 and 2010 exist in all models. 

* LME and SLR using data from the 2006-2010 period gives different predictions compared to other periods.

```{r fig.width = 8, fig.height = 8}
periods <- list.files("data/workingData", "lme_slr") %>% substr(., 9, 13)
pred_all <- lapply(paste0('data/workingData/', 
                          list.files("data/workingData", "lme_slr")), read.csv)

pred_all_clean <- lapply(pred_all, function(pred_df){
   pred_df <- pred_df %>% filter(year%in%(2009:2010)) %>% dplyr::select(slr, lme, year, df_type, period, obs)  #obs, period,
   names(pred_df)[1] <- paste0(names(pred_df)[1], '_', unique(pred_df$period))
   names(pred_df)[2] <- paste0(names(pred_df)[2], '_', unique(pred_df$period))
   pred_df
   # pred_df %>% select(-)
})
pred_all_clean <- do.call(cbind, pred_all_clean)
pred_all_clean <- pred_all_clean[, !duplicated(names(pred_all_clean))]

plot_df <- pred_all_clean %>% 
   filter(df_type=='test') %>% 
   dplyr::select(-year, -df_type, -period)
plot_df <- plot_df[,names(plot_df) %>% order]
pairs(plot_df, upper.panel=NULL)
```

# add year as a predictor
```{r}
tmp <- matrix(nrow=length(csv_names)*2, ncol=3)
tmp_rowname <- vector("character", length=length(csv_names)*2)
for(i in seq_along(years)){
   
   csv_name <- csv_names[i]
   csv_name
   no2_e_09_11 <- subset_df_yrs(no2_e_all, years[[i]])
   data_all <- no2_e_09_11
   #f# subset cross-validation data (5-fold cross-validation)
   #f# stratified by station types, climate zones and/or years
   set.seed(seed)
   data_all$index <- 1:nrow(data_all)
   # Test only leave location out first 
   # Method 1: (easier to use)
   folds=CreateSpacetimeFolds(
      data_all,
      spacevar = "station_european_code",     # leave location out
      timevar = NA,
      k = 5,
      class = NA,
      seed = seed
   )
   train_sub <- data_all[folds$index[[1]], ]
   test_sub <- data_all[folds$indexOut[[1]], ]
   source("scr/fun_call_predictor.R")
   #f# SLR: define/preprocess predictors (direction of effect)
   source("scr/fun_slr_proc_in_data.R")
   train_sub <- proc_in_data(train_sub, neg_pred)
   test_sub <- proc_in_data(test_sub, neg_pred)
   
   eq_lme <- as.formula(paste0('obs~',
                               paste(lme_result[[i]]$variables[-1], collapse = "+"),
                               "+ (1|station_european_code) + (1|year)"))
   lme_model <- lmer(eq_lme, train_sub)
   
   eq_slr <- as.formula(paste0('obs~',
                               paste(slr_result[[i]]$variables[-1], collapse = "+")))
   slr_model <- lm(eq_slr, train_sub)
   
   data_all2 <- rbind(train_sub %>% mutate(df_type="train"),
                      test_sub %>% mutate(df_type='test'))
   
   
   # the prediction will use the unconditional (population-level) values for data with previously unobserved levels (or NAs).
   source("scr/fun_gen_pred_lme.R")
   lme_prediction <- gen_pred_lme(lme_model, data_all2, "obs")
   source("scr/fun_gen_pred_df.R")
   slr_prediction <- gen_pred_df(slr_model, data_all2, "obs")
   
   tb1 <- error_matrix(with(lme_prediction, obs[df_type=='test']), with(lme_prediction, lme[df_type=='test']))[c(1,5,7)]
   # print(paste0("slr: test data ", substr(csv_names[[i]], 12, nchar(csv_names[[i]]))))
   tb2 <- error_matrix(with(slr_prediction, obs[df_type=='test']), with(slr_prediction, slr[df_type=='test']))[c(1,5,7)] 
   tmp[seq(1,by=2, length.out = length(csv_names))[i],] <- tb1
   tmp_rowname[seq(1,by=2, length.out = length(csv_names))[i]] <- paste0(substr(csv_names[i], 6, nchar(csv_names[i])), "_lme")
   tmp[seq(2,by=2, length.out = length(csv_names))[i],] <- tb2
   tmp_rowname[seq(2,by=2, length.out = length(csv_names))[i]] <- paste0(substr(csv_names[i], 6, nchar(csv_names[i])), "_slr")
   colnames(tmp) <- names(tb2)
}
rownames(tmp) <- tmp_rowname
tbl1 <- tmp %>% kbl

for (i in seq_along(1:(nrow(tmp)/2))) {
  tbl1 <- row_spec(tbl1, ((i-1)*2+1):(i*2), background = spec_color((nrow(tmp)/2):1)[i])
}
tbl1

print("add year as a predictor")

tmp <- matrix(nrow=length(csv_names)*2, ncol=3)
tmp_rowname <- vector("character", length=length(csv_names)*2)
for(i in seq_along(years)){

   eq_lme <- as.formula(paste0('obs~',
                               paste(lme_result[[i]]$variables[-1], collapse = "+"),
                               " + year + (1|station_european_code) + (1|year)"))
   lme_model <- lmer(eq_lme, train_sub)
   
   eq_slr <- as.formula(paste0('obs~',
                               paste(slr_result[[i]]$variables[-1], collapse = "+"),
                               " + year"))
   slr_model <- lm(eq_slr, train_sub)
   source("scr/fun_gen_pred_lme.R")
   lme_prediction <- gen_pred_lme(lme_model, data_all2, "obs")
   source("scr/fun_gen_pred_df.R")
   slr_prediction <- gen_pred_df(slr_model, data_all2, "obs")
   # plot(slr_prediction$slr, lme_prediction$lme)
   # 
   # plot(with(slr_prediction, slr[df_type=='test']), with(lme_prediction, lme[df_type=='test']))
   # plot(with(slr_prediction, slr[df_type=='train']), with(lme_prediction, lme[df_type=='train']))
   # plot(with(lme_prediction, obs[df_type=='train']), with(lme_prediction, lme[df_type=='train']))
   # plot(lme_prediction$lme, lme_prediction$obs)
   # # test data
   # plot(with(lme_prediction, lme[df_type=='test']), with(lme_prediction, obs[df_type=='test']))
   # plot(with(slr_prediction, slr[df_type=='test']), with(slr_prediction, obs[df_type=='test']))
   # print(paste0("lme: test data ", substr(csv_names[[i]], 12, nchar(csv_names[[i]]))))
   tb1 <- error_matrix(with(lme_prediction, obs[df_type=='test']), with(lme_prediction, lme[df_type=='test']))[c(1,5,7)]
   # print(paste0("slr: test data ", substr(csv_names[[i]], 12, nchar(csv_names[[i]]))))
   tb2 <- error_matrix(with(slr_prediction, obs[df_type=='test']), with(slr_prediction, slr[df_type=='test']))[c(1,5,7)] 
   tmp[seq(1,by=2, length.out = length(csv_names))[i],] <- tb1
   tmp_rowname[seq(1,by=2, length.out = length(csv_names))[i]] <- paste0(substr(csv_names[i], 12, nchar(csv_names[i])), "_lme")
   tmp[seq(2,by=2, length.out = length(csv_names))[i],] <- tb2
   tmp_rowname[seq(2,by=2, length.out = length(csv_names))[i]] <- paste0(substr(csv_names[i], 12, nchar(csv_names[i])), "_slr")
   colnames(tmp) <- names(tb2)
}
rownames(tmp) <- tmp_rowname
tbl1 <- tmp %>% kbl

for (i in seq_along(1:(nrow(tmp)/2))) {
  tbl1 <- row_spec(tbl1, ((i-1)*2+1):(i*2), background = spec_color((nrow(tmp)/2):1)[i])
}
tbl1
```

